---
title: "Gaussian Naive Bayes Model For Diabetics prediction"
author: "Anas"
date: "7/21/2024"
output: html_document
---

## Install the required packages.
```{r , echo=FALSE}
pacman::p_load(tidyverse, GGally, gtools, skimr, mice, caret, pROC, class,car,CMplot, caret, ggplot2, caTools, devtools, naivebayes,tidyverse, reshape2, ggpubr, skimr, rpart.plot, performanceEstimation, e1071, splitTools, cluster, factoextra, ggfortify)
```


## Load the Diabetics prediction dataset.
```{r , echo=FALSE}
diab <- read_csv("C:\\Users\\Anas\\Documents\\Workshops\\diabetes.csv")

```

## Convert the categorical data into numeric categories.
```{r}
diab$hypertension <- as.numeric(as.factor(diab$hypertension))
diab$heart_disease  <- as.numeric(as.factor(diab$heart_disease))
diab$gender <- as.numeric(as.factor(diab$gender))
diab$smoking_history <- as.numeric(as.factor(diab$smoking_history))
```


# Modeling.

## Train and test dataset splitting.
We will split the dataset by 70/30 in accordance to our target proportion.
```{r}
library(rsample)
set.seed(100)

library(caTools)

sample = sample.split(diab$bmi, SplitRatio = .7)
diab.train <- subset(diab, sample == TRUE)
diab.test  <- subset(diab, sample == FALSE)
diab.test_step  <- subset(diab, sample == FALSE)


train_x = diab.train[,1:8]
train_y = as.factor(diab.train$diabetes) 

test_x = diab.test[,1:8]
test_y = as.factor(diab.test$diabetes)
```


# Gaussian Naive Bayes - Training using cross validation approach
```{r Gaussian_Naive_Bayes, echo=FALSE}
# train_y <- (as.factor(train$diabetes))
train_x1 <- as.matrix(train_x)
test_x1  <- as.matrix(test_x)

library(naivebayes)
GNB = gaussian_naive_bayes(train_x1, train_y)

pred_GNB_DB = predict(GNB, newdata = test_x1, type = "prob")

summary(GNB)
GNB$params
GNB$prior
coef(GNB)
```

## Gaussian Naive Bayes prediction results
```{r}
# Convert data frame to matrix as required by the model. 
test_x = as.matrix(sapply(test_x, as.numeric))

# Perform predictions for the test data. Type can be class or prob for GNB.
pred = predict(GNB, newdata = test_x, type = "class")

library(caret)
GNB_comf_mat <- confusionMatrix(data=test_y, reference=pred, positive="1")
GNB_comf_mat

#length(test_y)
#length(pred)

```

# Work for Brier Score - IMBALANCDED dataset

```{r}
pred_DB = predict(GNB, newdata = test_x1, type = "prob")
test_y = as.numeric(test_y)

sqdiff = (test_y - pred_DB)^2
sqdiff.rm.na = complete.cases(sqdiff)  # remove rows with NA values.

BrierScore= mean(sqdiff.rm.na)
BrierScore
```

# ---------------------------------------------------
# SMOTE on the training data and the GNB model on it. 
# ---------------------------------------------------
```{r}

library(DMwR2)
set.seed(12345)
diab_balanced = smote(diabetes ~ ., data = diab, perc.over = 1)

library(rsample)
set.seed(100)

library(caTools)

smp_size <- floor(0.70 * nrow(diab_balanced))
set.seed(123)
train_ind  <- sample(seq_len(nrow(diab_balanced)), size=smp_size)

train <- diab_balanced[train_ind, ]
test  <- diab_balanced[-train_ind,]

xtrain_smote <- train[, 1:8] 
ytrain_smote <- as.factor(train$diabetes)
  
xtest_smote <- test[, 1:8]
ytest_smote <- as.factor(test$diabetes)

##
library(smotefamily)
ytrain_smote_GNB <- smote(diabetes ~ ., data = diab.train, perc.over = 200, perc.under = 100)


```

# ----------------------------------------------------------------------
# Gaussian Naive Bayes Classification with model validation on test data 
# ----------------------------------------------------------------------

```{r}

library(naivebayes)
xtrain_smote1 = as.matrix(xtrain_smote)

GNB_SMOTE = gaussian_naive_bayes(xtrain_smote1, ytrain_smote)
summary(GNB_SMOTE) 
GNB_SMOTE$params
GNB_SMOTE$prior
coef(GNB_SMOTE)

```

# Predictions using SMOTE based dataset
```{r}
# Convert data frame to matrix as requried by the model. 
xtest_smote1 = as.matrix(sapply(xtest_smote, as.numeric))


# Perform predictions for the test data. Type can be class or prob for GNB.
pred_DB = predict(GNB_SMOTE, newdata = xtest_smote1, type = "class")

library(caret)
GNB_comf_mat_SMOTE <- confusionMatrix(data = ytest_smote, reference = pred_DB, positive = "1")
GNB_comf_mat_SMOTE

```

# Work for Brier Score - Balanced dataset
```{r}
pred_DB = predict(GNB_SMOTE, newdata = xtest_smote1, type = "prob")

ytest_smote = as.numeric(ytest_smote)

sqdiff = (ytest_smote - pred_DB)^2
sqdiff.rm.na = complete.cases(sqdiff)  # remove rows with NA values.

BrierScore= mean(sqdiff.rm.na)
BrierScore
```