---
title: "Logistic Regression For Diabetics prediction"
author: "Anas"
date: "8/31/2024"
output: html_document
---

# Install the required packages.
```{r }
pacman::p_load(tidyverse, GGally, gtools, skimr, mice, caret, pROC, class,car,CMplot, caret, ggplot2, caTools, devtools, naivebayes,tidyverse, reshape2, ggpubr, skimr, rpart.plot, performanceEstimation, e1071, splitTools, cluster, factoextra, ggfortify)
```


## Load the Diabetics prediction dataset.
```{r , echo=FALSE}
diab <- read.csv ("C:\\Users\\Anas\\Documents\\Workshops\\diabetes.csv")


```

```{r}
# skim(diab)
# boxplot(diab, las = 2, varwidth = T)
# ggcorr(diab, label = T)
```


```{r}
diab$hypertension <- as.numeric(as.factor(diab$hypertension))
diab$heart_disease  <- as.numeric(as.factor(diab$heart_disease))
diab$gender <- as.numeric(as.factor(diab$gender))
diab$smoking_history <- as.numeric(as.factor(diab$smoking_history))

```


```{r}
#prop.table(table(diab$diabetes))
```


# Modeling.
## Train and test dataset splitting.
We will split the dataset by 70/30 in accordance to our target proportion.

```{r}
library(rsample)
set.seed(100)

require(caTools)

sample = sample.split(diab$bmi, SplitRatio = .7)
diab.train <- subset(diab, sample == TRUE)
diab.test  <- subset(diab, sample == FALSE)
```


# EDA on train dataset
```{r}
#skim(diab.train)
#boxplot(diab.train, las = 2, varwidth = T)
#ggcorr(diab.train, label = T)
```

# Logistic Regression on imbalanced data
```{r}
model1 <- glm( diabetes ~ .,family = "binomial", data = diab.train)
summary(model1)

# Paramaeters' Interpretation
exp(model1$coefficients)

```

# Model Evaluation
## Residual deviance, AIC and VIF of each models
```{r}
model1$deviance
model1$aic

# VIF: #vif(model1)

```


# LR model Prediction for imbalanced data
# ----------------------------------------
```{r}
# Using 0.3 as the threshold probability.
# --------------------------------------
diab.test$pred.Outcome <- predict(model1,newdata=diab.test,type = "response")
prob = diab.test$pred.Outcome
diab.test$pred.Outcome <-as.factor(ifelse(diab.test$pred.Outcome > 0.7, "1","0"))

```


# -------------------------------------------------------------
# Model Comparison using confusion matrix eith imbalanced data
# -------------------------------------------------------------
```{r}
predicted_y      <- diab.test$pred.Outcome
test_y <- (as.factor(diab.test$diabetes))
confusionMatrix(data = test_y,reference = predicted_y, positive = "1")


# Brier Score for Imbalanced data
# -------------------------------
diab.test$diabetes <- as.numeric(diab.test$diabetes)
sqdiff = (diab.test$diabetes - prob)^2
#sqdiff.rm.na = complete.cases(sqdiff)  # remove rows with NA values.

BrierScore= mean(sqdiff)
BrierScore



```


# ------------------------------------------------------------------------------
#                    SMOTE over-sampling from minorty class
# ------------------------------------------------------------------------------
Since our outcome is not balanced, we need to balance the data before we make a model out of it. using SMOTE.


# ----------------------------------------------------------
# Logistic Regression with Brier Score with model validation
# ----------------------------------------------------------
```{r}
set.seed(12345)
train.balanced <- smote(diabetes ~ ., data = diab, perc.over = 1)

library(rsample); set.seed(100)
sample = sample.split(train.balanced$bmi, SplitRatio = .7)
train <- subset(train.balanced, sample == TRUE)
test  <- subset(train.balanced, sample == FALSE)
```


# Logistic Regression for balanced data
# -------------------------------------
```{r}
train$diabetes <- as.factor(train$diabetes)
mod_SMOTE <- glm( diabetes ~ .,family = "binomial", data = train)
pred_diab <- predict(mod_SMOTE, test, type = "response")

# Store the probabilities in prob
test$prob= pred_diab;

# Segmentation for Classification task.
test$pred_diab <- factor(ifelse(pred_diab > 0.5, 1,0))

```


# Brier Score for LR model using Imbalanced data
# ---------------------------------------------
```{r}
predicted_y      <- test$pred_diab
test_y <- (as.factor(test$diabetes))
                     
confusionMatrix(data = test_y, reference = predicted_y, positive = "1")


test$diabetes <- as.numeric(test$diabetes)
sqdiff = (test$diabetes - test$prob)^2
sqdiff.rm.na = complete.cases(sqdiff)  # remove rows with NA values.

BrierScore= mean(sqdiff.rm.na)
BrierScore
```
