---
title: "Non-Parametric Naive Bayes Model For Diabetics prediction"
author: "Anas"
output: html_document
---

## Install the required packages.
```{r , echo=FALSE}
pacman::p_load(tidyverse, GGally, gtools, skimr, mice, caret, pROC, class,car,CMplot, caret, ggplot2, caTools, devtools, naivebayes,tidyverse, reshape2, ggpubr, skimr, rpart.plot, performanceEstimation, e1071, splitTools, cluster, factoextra, ggfortify)
```


## Load the Diabetics prediction dataset.
```{r , echo=FALSE}
diab <- read.csv("C:\\Users\\Anas\\Documents\\Workshops\\diabetes.csv")

```


## Convert the categorical data into numeric categories.
```{r}
diab$hypertension <- as.numeric(as.factor(diab$hypertension))
diab$heart_disease  <- as.numeric(as.factor(diab$heart_disease))
diab$gender <- as.numeric(as.factor(diab$gender))
diab$smoking_history <- as.numeric(as.factor(diab$smoking_history))

```


# Modeling.
## Train and test dataset splitting.
We will split the dataset by 70/30 in accordance to our target proportion.
```{r}
library(rsample)
set.seed(100)

library(caTools)

sample = sample.split(diab$bmi, SplitRatio = .7)
diab.train <- subset(diab, sample == TRUE)
diab.test  <- subset(diab, sample == FALSE)
diab.test_step  <- subset(diab, sample == FALSE)


train_x = diab.train[,1:8]
train_y = as.factor(diab.train$diabetes) 

test_x = diab.test[,1:8]
test_y = as.factor(diab.test$diabetes)
```



# Non-Parametric Naive Bayes - Training using cross validation approach
```{r, echo=FALSE}
library(naivebayes)

train_x1 <- as.matrix(train_x)

NPNB <- nonparametric_naive_bayes(x = train_x1, y = train_y)
NPNB
summary(NPNB)
NPNB$params
NPNB$prior
coef(NPNB)
```


## Non-Parametric Naive Bayes prediction results
```{r}

# Convert data frame to matrix as requried by the model. 
test_x1= as.matrix(sapply(test_x, as.numeric))


# Perform predictions for the test data. Type can be class or prob for NPNB.
pred_prob <- predict(NPNB, newdata = test_x1, type = "prob")
pred      <- predict(NPNB, newdata = test_x1, type = "class")


library(caret)
NPNB_comf_mat <- confusionMatrix(data = test_y, reference = pred, positive = "1")
NPNB_comf_mat

```

# Model Interpretation
To do

########################################################################

# SMOTE on the training data and the NPNB model on it. 
```{r}

library(DMwR2) # this is obsolete package
#install.packages("smotefamily")
#library(smotefamily)
#diab_balanced = SLS(diab[,-9],diab[,9] , C=5)

set.seed(12345)
diab_balanced = smote(diabetes ~ ., data = diab, perc.over = 1)


library(rsample)
set.seed(100)

library(caTools)

smp_size <- floor(0.70 * nrow(diab_balanced))
set.seed(123)
train_ind  <- sample(seq_len(nrow(diab_balanced)), size=smp_size)

train <- diab_balanced[train_ind, ]
test  <- diab_balanced[-train_ind,]

xtrain_smote <- train[, 1:8] 
ytrain_smote <- as.factor(train$diabetes)
  
xtest_smote <- test[, 1:8]
ytest_smote <- as.factor(test$diabetes)


```


```{r}

library(naivebayes)
xtrain_smote1 = as.matrix(xtrain_smote)

NPNB_SMOTE = nonparametric_naive_bayes(xtrain_smote1, ytrain_smote)


summary(NPNB_SMOTE) 
NPNB_SMOTE$params
NPNB_SMOTE$prior
coef(NPNB_SMOTE)

```

# Predictions using SMOTE based dataset
```{r}
# Convert data frame to matrix as requried by the model. 
xtest_smote1 = as.matrix(sapply(xtest_smote, as.numeric))


# Perform predictions for the test data. Type can be class or prob for NPNB.
pred_DB = predict(NPNB_SMOTE, newdata = xtest_smote1, type = "class")

library(caret)
NPNB_comf_mat_SMOTE <- confusionMatrix(data = ytest_smote, reference = pred_DB, positive = "1")
NPNB_comf_mat_SMOTE

```

